---
title: "Data Cleaning Project in R using Iris Data"
author: "Melanie Gruben"
date: "2022-11-02"
output: html_document
---

### In this project I will use R to clean dirty data from a dataset of iris flower specimens.

First I'll set the ***working directory***:

```{r}
setwd("/cloud/project/Data Cleaning")
```

Next I'll label the data and look over it.

```{r}
data <- read.csv("iris_dirty.csv")
head(data)
```
We can see right away that some of the capitalisation is inconsistent in the species names. We'll get to that in a bit.


```{r}
install.packages("janitor")
library(janitor)
```

Let's now ***make all the variable names consistent*** and relabel the new data.

```{r}
data2 <-clean_names(data)
```

Now let's view the effects of the cleaning.

```{r}
data2
```

Perfect. There may also be ***empty rows or columns***; let's check for and remove those.

```{r}
data3 <- remove_empty(data2, which = c("rows", "cols"), quiet = FALSE)
```
No issues there, but it was a good measure. Now let's use dplyr to ***clear out any duplicate data***.

```{r}
library(dplyr)
data_cleaned <- distinct(data3)
data_cleaned
```

We can also clear duplicate data with regards to a ***specific variable or column***:

```{r}
data_cleaned2 <- distinct(data_cleaned, petal_length, .keep_all = TRUE)
data_cleaned2
```

And finally let's ***correct the capitalisation*** in the species column to make it consistently lowercase, just because it's nicer looking that way.

```{r}
data_cleaned2 %>% mutate(species = tolower(species))
```

***And there we have it***; we've done a quick clean of dirty data in R using janitor and dplyr. Thanks for reading!

-Melanie
